Stan Practice

### 2021-02-16 Stan/Stan Base Models.Rmd
Mixed Models (With Ant Interaction)

```{r}
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}
```

## Basic Models (No Ant Effects)
```{r Growth Model}
## Now rename variables and index
x <- log(data$volume_t)
z <- as.numeric(data$ant_t)
y <- log(data$volume_t1)
N <- length(data$volume_t)
K <- length(data$ant_t)
## Rerun linear model with that new data
#mod1 <- lm(y~x, data = data)
#summary(grow_model)
#summary(mod1)
## Now we can extract some important statistics
lm_alpha <- summary(grow_model)$coeff[1]  # the intercept
lm_beta <- summary(grow_model)$coeff[2]  # the slope
lm_sigma <- sigma(grow_model)  # the residual error
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y, z = z, K = K)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression

data {
 int < lower = 1 > N; // Sample size
 vector[N] x; // Predictor
 vector[N] y; // Outcome
}

@@

model {
// Priors
alpha ~ normal(0,100);
beta ~ normal(0,100); 
//sigma ~ exponential(0,0);
 y ~ normal(alpha + x * beta , sigma); // The Model
}

@@

"grow_lm1.stan")
## Check that you wrote a file
#stanc("grow_lm1.stan")
## Save the file path
stanny <- "grow_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit
#class(fit)
posterior <- as.data.frame(fit)
print(colnames(posterior))

par(mfrow = c(1,3))
plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
plot(density(posterior$beta), main = "Beta")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

```{r Total Flowers}
x <- log(data$volume_t)
y <- data$TotFlowerbuds_t
N <- length(data$volume_t)
## Rerun linear model with that new data
#mod1 <- lm(y~x, data = data)
summary(flower_model)
#summary(mod1)
## Now we can extract some important statistics
lm_lambda = mean(data$TotFlowerbuds_t)
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple total flowers regression

data {
 int N;
 real x[N];
 int y[N];
}
parameters {
 real<lower = 0> lambda;
}
model {
// Priors
 lambda ~ cauchy(0, 1);
 //lambda ~ gamma(0, 0);
 // Model
 y ~ poisson(lambda);
}
",

"flower_lm1.stan")
## Check that you wrote a file
stanc("flower_lm1.stan")
## Save the file path
stanny <- "flower_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

plot(density(posterior$lambda), main = "Lambda")
abline(v = lm_lambda, col = 4, lty = 2)
```

```{r Survival}
x <- log(data$volume_t)
y <- data$Survival_t1
N <- length(data$volume_t)
#import things from the r regression models for comparison
lm_alpha <- summary(surv_model)$coeff[1]  # the intercept
lm_beta <- summary(surv_model)$coeff[2]  # the slope
summary(surv_model)
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple survival regression

data {
  int<lower=0> N;
  vector[N] x;
  int<lower=0,upper=1> y[N];
}
parameters {
  real alpha;
  real beta;
}
model {
 // Priors
 alpha ~ normal(0,100);
 beta ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(alpha + beta * x);
}
",

"surv_lm1.stan")
## Check that you wrote a file
stanc("surv_lm1.stan")
## Save the file path
stanny <- "surv_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

@@

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)

plot(density(posterior$beta), main = "Beta")
abline(v = lm_beta, col = 4, lty = 2)
```

#### Include ants
######Growth
```{r Growth Model}
## Now rename variables and index
vol <- log(data$volume_t)
ant <- as.numeric(data$ant_t)
y <- log(data$volume_t1)
N <- length(data$volume_t)
K <- length(data$ant_t)
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
lm_alpha <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
lm_beta <- summary(grow_model_ant_interact)$coeff[2]  # the slope
lm_sigma <- sigma(grow_model_ant_interact)  # the residual error

write("// Stan model for ant growth regression

data {
 int < lower = 1 > N; // Sample size
 int < lower = 1 > K; // Number of Predictors (Ant Species)
 real < lower = 0 > vol[N]; // volume
 int < lower = 0 > ant[K]; // ants
 real < lower = 0 > y[N]; // outcome vector
}

parameters {
 real alpha; // Intercept
 vector[K] beta; // coefficients for predictors
 real < lower = 0 > sigma; // Error SD
}

transformed parameters{
real mu[N];
for(i in 1:N){
mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
             }
}

model {
// Priors
alpha ~ normal(0,100);
beta ~ normal(0,100); 
//sigma ~ exponential(0,0);
 y ~  normal(mu, sigma);   // The Model
}

"grow_ant_lm1.stan")
## Check that you wrote a file
stanc("grow_ant_lm1.stan")
## Save the file path
stanny <- "grow_ant_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000 , chains = 4, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)

#plot(density(posterior$beta[i]), main = "Beta")
#abline(v = lm_beta, col = 4, lty = 2)

plot(density(posterior$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

```{r grow attempt 1}
vol = data$volume_t
y = data$volume_t1
crem = data$crem
liom = data$liom
vacant = data$vacant
other = data$other
N = nrow(data)
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  crem = crem, liom = liom, vacant = vacant, other = other) ## predictors ants

write("
      data{
      int <lower = 1> N; // Number of observations
      vector[N] vol; // volume predictor
      int <lower = 0, upper = 1> crem[N]; // crem predictor
      int <lower = 0, upper = 1> liom[N]; // liom predictor
      int <lower = 0, upper = 1> other[N]; // other predictor
      int <lower = 0, upper = 1> vacant[N]; // vacant predictor
      vector[N] y; // outcome
      }
      parameters{
      real alpha; // intercept 
      real b_crem; // crem slope
      real b_liom; // liom slope
      real b_other; // other slope
      real b_vacant; // vacant slope
      real <lower = 0> sigma; // standard deviation
      }
      model{
      // Priors
      alpha ~ normal(0,100);
      beta ~ normal(0,100);
      // Likelihood
      y ~ normal(alpha + (b_crem*crem[N] + b_liom*liom[N] + b_other*other[N] + b_vacant*vacant[N])*vol ,sigma);
      }
      ",
      "tester1.stan")

stanc("tester1.stan")

fit <- stan(file = "tester1.stan", data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))
par(mfrow = c(2,3))
plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
plot(density(posterior$b_other), main = "Beta Other")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$b_vacant), main = "Beta Vacant")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$b_crem), main = "Beta Crem")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$b_liom), main = "Beta Liom")
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

```{r grow attempt 2&3}
vol <- log(data$volume_t)
ant <- as.numeric(data$ant_t)
y <- log(data$Survival_t1)
N <- length(data$volume_t)
K <- length(data$ant_t)
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
## Write the model in stan
lm_alpha <- summary(surv_model_ant)$coeff[1]  # the intercept
lm_beta <- summary(surv_model_ant)$coeff[2]  # the slope
lm_sigma <- sigma(surv_model_ant_interact)  # the residual error
write("// Stan model for simple linear regression

data {
 int < lower = 1 > N; // Sample size
 int < lower = 1 > K; // Number of Predictors (Ant Species)
 real < lower = 0 > vol[N]; // volume
 int < lower = 0 > ant[K]; // ants
 int <lower=0,upper=1> y[N]; // outcome vector
}

parameters {
  real alpha;
  vector[K] beta; // coefficients for predictors
}

transformed parameters{
 real mu[N];
 for(i in 1:N){
 mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
 }
}

model {
//Priors
 alpha ~ normal(0,100);
 beta ~ normal(0,100);
 //Model
 y ~ bernoulli_logit(mu);
}
",

"surv_ant_lm1.stan")
## Check that you wrote a file
stanc("surv_ant_lm1.stan")
## Save the file path
stanny <- "surv_ant_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 5000, iter = 10000, chains = 1, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
```

######Survival

```{r survival}
vol <- log(data$volume_t)
ant <- as.numeric(data$ant_t)
y <- log(data$Survival_t1)
N <- length(data$volume_t)
K <- length(data$ant_t)
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y, ant = ant, K = K)
## Write the model in stan
lm_alpha <- summary(surv_model_ant)$coeff[1]  # the intercept
lm_beta <- summary(surv_model_ant)$coeff[2]  # the slope
lm_sigma <- sigma(surv_model_ant_interact)  # the residual error
write("// Stan model for simple ant survival regression

data {
 int < lower = 1 > N; // Sample size
 int < lower = 1 > K; // Number of Predictors (Ant Species)
 real < lower = 0 > vol[N]; // volume
 int < lower = 0 > ant[K]; // ants
 int <lower=0,upper=1> y[N]; // outcome vector
}

parameters {
  real alpha;
  vector[K] beta; // coefficients for predictors
}

transformed parameters{
 real mu[N];
 for(i in 1:N){
 mu[i] = beta[1] + beta[2]*vol[i] + beta[3]*ant[i]; 
 }
}

model {
//Priors
 alpha ~ normal(0,100);
 beta ~ normal(0,100);
 //Model
 y ~ bernoulli_logit(mu);
}
",

"surv_ant_lm1.stan")
## Check that you wrote a file
stanc("surv_ant_lm1.stan")
## Save the file path
stanny <- "surv_ant_lm1.stan"
## Now we fit the model
fit <- stan(file = stanny, data = stan_data, warmup = 5000, iter = 10000, chains = 1, cores = 2, thin = 1)
fit

posterior <- as.data.frame(fit)
print(colnames(posterior))

plot(density(posterior$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
```

```{r surv attempt 1}
vol = data$volume_t
y = as.integer(data$Survival_t1)
crem = data$crem
liom = data$liom
vacant = data$vacant
other = data$other
N = nrow(data)
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  crem = crem, liom = liom, vacant = vacant, other = other) ## predictors ants

write("
      data{
      int <lower = 1> N; // Number of observations
      vector[N] vol; // volume predictor
      int <lower = 0, upper = 1> crem[N]; // crem predictor
      int <lower = 0, upper = 1> liom[N]; // liom predictor
      int <lower = 0, upper = 1> other[N]; // other predictor
      int <lower = 0, upper = 1> vacant[N]; // vacant predictor
      int <lower = 0, upper = 1> y[N]; // outcome
      }
      parameters{
      real alpha; // intercept 
      real b_crem; // crem slope
      real b_liom; // liom slope
      real b_other; // other slope
      real b_vacant; // vacant slope
      real <lower = 0> sigma; // standard deviation
      }
      model{
      // Priors
      alpha ~ normal(0,100);
      // Likelihood
      y ~ bernoulli_logit(alpha + (b_crem*crem[N] + b_liom*liom[N] + b_other*other[N] + b_vacant*vacant[N])*vol);
      }
      ",
      "tester1.stan")

stanc("tester1.stan")

fit <- stan(file = "tester1.stan", data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit
```

>>>>

```{r}
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1","repro_state_t1")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}

flower <- cactus[ , c("TotFlowerbuds_t","ABFlowerbuds_t","Goodbuds_t","ant_t", "repro_state_t1","volume_t","volume_t1")]
flower <- na.omit(flower)
flower$ant <- as.integer(flower$ant_t)
```

## Basic Models (No Ant Effects)
####Growth
```{r Growth Model}
## Now rename variables and index
vol <- data$volume_t
y <- data$volume_t1
N <- as.integer(nrow(data))
## Rerun linear model with that new data
#mod1 <- lm(y~x, data = data)
#summary(grow_model)
#summary(mod1)
## Now we can extract some important statistics
lm_alpha <- summary(grow_model)$coeff[1]  # the intercept
lm_beta <- summary(grow_model)$coeff[2]  # the slope
lm_sigma <- sigma(grow_model)  # the residual error
## Make it a dataframe
stan_data <- list(N = N, vol = vol, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression

data {
 int < lower = 1 > N; // Sample size
 vector[N] vol; // Predictor
 vector[N] y; // Outcome
}

@@

model {
// Priors
alpha ~ normal(0,100);
beta ~ normal(0,100); 
//sigma ~ exponential(0,0);
 y ~ normal(alpha + vol * beta , sigma); // The Model
}

@@

"grow_lm1.stan")
## Check that you wrote a file
#stanc("grow_lm1.stan")
## Save the file path
stanny <- "grow_lm1.stan"
## Now we fit the model
fit_grow <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_grow
#class(fit)
posterior_grow <- as.data.frame(fit_grow)
print(colnames(posterior_grow))

par(mfrow = c(1,3))
plot(density(posterior_grow$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)
plot(density(posterior_grow$beta), main = "Beta", xlim = c(0.85,0.93))
abline(v = lm_beta, col = 4, lty = 2)
plot(density(posterior_grow$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

####Total Flowers
```{r Total Flowers}
## Ok so this one is difficult because the data does not include this variable because of all of the NAs. 
## Ok so I have created a second data set. Obviously this will be a problem down the line when it all has to go into one model, but here we go for now. 
## Next problem is that the dimensions do not work now. 
y <- flower$TotFlowerbuds_t
N <- nrow(flower)
stan_data <- list(N = N, ## number of observations
                  y = y ## response survival next year
                  )

write("// Stan model for simple total flowers regression

data {
 int < lower = 1 > N; // Sample size
 int y[N]; // Outcome
}
parameters {
  real lambda; //intercept, unique to ant
}
model {
// Priors
 lambda ~ beta(1,1);
 // Model
 y ~ poisson(lambda);
}
",

"flower_lm1.stan")
## Check that you wrote a file
stanc("flower_lm1.stan")
## Save the file path
stanny <- "flower_lm1.stan"
## Now we fit the model
fit_flow <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_flow

posterior_flow <- as.data.frame(fit_flow)
print(colnames(posterior_flow))

plot(density(posterior_flow$lambda), main = "Lambda")
abline(v = lm_lambda, col = 4, lty = 2)
```
####Abort
```{r abort model}
y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
vol = flower$volume_t
N = nrow(flower)
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol = vol
                  )

write("// Stan model for simple total flowers regression

data {
 int < lower = 1 > N; // Sample size
 int <lower = 0, upper = 1> y[N]; // Outcome
 vector[N] vol;	//size_t
}
parameters {
  real beta0; //intercept, unique to ant sp
  real beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
model {
 // Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(beta0 + beta1 * vol);
}
",

"abort_lm1.stan")
## Check that you wrote a file
stanc("abort_lm1.stan")
## Save the file path
stanny <- "abort_lm1.stan"
## Now we fit the model
fit_abort <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_abort

posterior_abort <- as.data.frame(fit_abort)
print(colnames(posterior_abort))
```

####Survival
```{r Survival}
vol = data$volume_t
y <- data$Survival_t1
N <- nrow(data)
#import things from the r regression models for comparison
lm_alpha <- summary(surv_model)$coeff[1]  # the intercept
lm_beta <- summary(surv_model)$coeff[2]  # the slope
summary(surv_model)
## Make it a dataframe
stan_data <- list(N = N, x = x, y = y)
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple survival regression

data {
  int <lower=0> N;
  vector[N] vol;
  int <lower=0,upper=1> y[N];
}
parameters {
  real alpha;
  real beta;
}
model {
 // Priors
 alpha ~ normal(0,100);
 beta ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(alpha + beta * vol);
}
",

"surv_lm1.stan")
## Check that you wrote a file
stanc("surv_lm1.stan")
## Save the file path
stanny <- "surv_lm1.stan"
## Now we fit the model
fit_surv <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_surv

posterior_surv <- as.data.frame(fit_surv)
print(colnames(posterior_surv))

@@

plot(density(posterior_surv$alpha), main = "Alpha")
abline(v = lm_alpha, col = 4, lty = 2)

plot(density(posterior_surv$beta), main = "Beta")
abline(v = lm_beta, col = 4, lty = 2)
```

####Repro

```{r}
vol1 <- flower$volume_t1
y = flower$repro_state_t1
N = nrow(flower)
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol1 = vol1
                  )

write("// Stan model for simple total flowers regression

data {
 int < lower = 1 > N; // Sample size
 int <lower = 0, upper = 1> y[N]; // Outcome
 vector[N] vol1;	//size_t
}
parameters {
  real beta0; //intercept, unique to ant sp
  real beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
model {
 // Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(beta0 + beta1 * vol1);
}
",

"repro_lm1.stan")

## Check that you wrote a file
stanc("repro_lm1.stan")
## Save the file path
stanny <- "repro_lm1.stan"
## Now we fit the model
fit_repro <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_repro

posterior_repro <- as.data.frame(fit_repro)
print(colnames(posterior_repro))
```

## Include ants
####Growth
```{r Growth Model}
vol = data$volume_t
y = data$volume_t1
N = nrow(data)
N_ant = 4
ant = data$ant
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant
                  ) 
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  vector[N] y; // size_t1
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
  }
}

model {
  y ~ normal(mu, sigma); // likelihood
}",

"grow_ant_lm1.stan")
## Check that you wrote a file
stanc("grow_ant_lm1.stan")
## Save the file path
#stanny <- "grow_ant_lm1.stan"
## Now we fit the model
fit_grow_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_grow_ant
#class(fit)
posterior_grow_ant <- as.data.frame(fit_grow_ant)
print(colnames(posterior_grow_ant))
lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(grow_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(grow_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(grow_model_ant_interact)$coeff[5]
lm_beta_other <- summary(grow_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(grow_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(grow_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(grow_model_ant_interact)$coeff[8]
lm_sigma <- sigma(grow_model_ant_interact)  # the residual error

par(mfrow = c(2,4))
#Plot Intercepts
plot(density(posterior_grow_ant$`beta0[1]`), main = "Intercept Other")
abline(v = lm_alpha_other, col = 4, lty = 2)
plot(density(posterior_grow_ant$`beta0[2]`), main = "Intercept Crem")
abline(v = lm_alpha_crem, col = 4, lty = 2)
plot(density(posterior_grow_ant$`beta0[3]`), main = "Intercept Liom")
abline(v = lm_alpha_liom, col = 4, lty = 2)
plot(density(posterior_grow_ant$`beta0[4]`), main = "Intercept Vacant")
abline(v = lm_alpha_vacant, col = 4, lty = 2)
#Plot Slopes
plot(density(posterior_grow_ant$`beta1[1]`), main = "Intercept Other")
abline(v = lm_beta_other, col = 4, lty = 2)
plot(density(posterior_grow_ant$`beta1[2]`), main = "Intercept Crem")
abline(v = lm_beta_crem, col = 4, lty = 2)
plot(density(posterior_grow_ant$`beta1[3]`), main = "Intercept Liom")
abline(v = lm_beta_liom, col = 4, lty = 2)
plot(density(posterior_grow_ant$`beta1[4]`), main = "Intercept Vacant")
abline(v = lm_beta_vacant, col = 4, lty = 2)
#Sigma
plot(density(posterior_grow_ant$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

####Survival

```{r survival}
vol = data$volume_t
y = data$Survival_t1
N = nrow(data)
N_ant = 4
ant = data$ant
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response survival next year
                  ant = ant,## predictors ants
                  N_ant = N_ant
                  )

write("// Stan model for simple ant survival regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
}

parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
  }
}

model {
//Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100);
 //Model
 y ~ bernoulli_logit(mu);
}
",

"surv_ant_lm1.stan")
## Check that you wrote a file
stanc("surv_ant_lm1.stan")
## Save the file path
stanny <- "surv_ant_lm1.stan"
## Now we fit the model
fit_surv_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_surv_ant

#class(fit)
posterior_surv_ant <- as.data.frame(fit_surv_ant)
print(colnames(posterior_surv_ant))
lm_alpha_other <- summary(surv_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(surv_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(surv_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(surv_model_ant_interact)$coeff[5]
lm_beta_other <- summary(surv_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(surv_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(surv_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(surv_model_ant_interact)$coeff[8]
lm_sigma <- sigma(surv_model_ant_interact)  # the residual error

par(mfrow = c(2,4))
#Plot Intercepts
plot(density(posterior_surv_ant$`beta0[1]`), main = "Intercept Other")
abline(v = lm_alpha_other, col = 4, lty = 2)
plot(density(posterior_surv_ant$`beta0[2]`), main = "Intercept Crem")
abline(v = lm_alpha_crem, col = 4, lty = 2)
plot(density(posterior_surv_ant$`beta0[3]`), main = "Intercept Liom")
abline(v = lm_alpha_liom, col = 4, lty = 2)
plot(density(posterior_surv_ant$`beta0[4]`), main = "Intercept Vacant")
abline(v = lm_alpha_vacant, col = 4, lty = 2)
#Plot Slopes
plot(density(posterior_surv_ant$`beta1[1]`), main = "Intercept Other")
abline(v = lm_beta_other, col = 4, lty = 2)
plot(density(posterior_surv_ant$`beta1[2]`), main = "Intercept Crem")
abline(v = lm_beta_crem, col = 4, lty = 2)
plot(density(posterior_surv_ant$`beta1[3]`), main = "Intercept Liom")
abline(v = lm_beta_liom, col = 4, lty = 2)
plot(density(posterior_surv_ant$`beta1[4]`), main = "Intercept Vacant")
abline(v = lm_beta_vacant, col = 4, lty = 2)
#Sigma
plot(density(posterior_surv_ant$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

####Total Flowers

```{r total flowers}
y <- flower$TotFlowerbuds_t
vol = flower$volume_t
N = nrow(flower)
N_ant = 4
ant = flower$ant
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol = vol,
                  N_ant = N_ant,
                  ant = ant
)

write("// Stan model for simple total flowers regression

data {
 int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0> y[N]; // survival in year t1
}

parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
  }
}
model {
 // Model
 y ~ poisson(mu);
}
",

"flower_lm1.stan")
## Check that you wrote a file
stanc("flower_lm1.stan")
## Save the file path
stanny <- "flower_lm1.stan"
## Now we fit the model
fit_flow_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_flow_ant

posterior_flow_ant <- as.data.frame(fit_flow_ant)
print(colnames(posterior_flow_ant))

lm_alpha_other <- summary(flower_model_ant)$coeff[1]  # the intercept
lm_alpha_crem <- summary(flower_model_ant)$coeff[3]
lm_alpha_liom <- summary(flower_model_ant)$coeff[4]
lm_alpha_vacant <- summary(flower_model_ant)$coeff[5]
lm_beta_other <- summary(flower_model_ant)$coeff[2]  # the slope
lm_beta_crem <- summary(flower_model_ant)$coeff[6]
lm_beta_liom <- summary(flower_model_ant)$coeff[7]
lm_beta_vacant <- summary(flower_model_ant)$coeff[8]

par(mfrow = c(2,4))
#Plot Intercepts
plot(density(posterior_flow_ant$`beta0[1]`), main = "Intercept Other")
abline(v = lm_alpha_other, col = 4, lty = 2)
plot(density(posterior_flow_ant$`beta0[2]`), main = "Intercept Crem")
abline(v = lm_alpha_crem, col = 4, lty = 2)
plot(density(posterior_flow_ant$`beta0[3]`), main = "Intercept Liom")
abline(v = lm_alpha_liom, col = 4, lty = 2)
plot(density(posterior_flow_ant$`beta0[4]`), main = "Intercept Vacant")
abline(v = lm_alpha_vacant, col = 4, lty = 2)
#Plot Slopes
plot(density(posterior_flow_ant$`beta1[1]`), main = "Intercept Other")
abline(v = lm_beta_other, col = 4, lty = 2)
plot(density(posterior_flow_ant$`beta1[2]`), main = "Intercept Crem")
abline(v = lm_beta_crem, col = 4, lty = 2)
plot(density(posterior_flow_ant$`beta1[3]`), main = "Intercept Liom")
abline(v = lm_beta_liom, col = 4, lty = 2)
plot(density(posterior_flow_ant$`beta1[4]`), main = "Intercept Vacant")
abline(v = lm_beta_vacant, col = 4, lty = 2)
```

####Abortion Model

```{r abortion model}
y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
vol = flower$volume_t
N = nrow(flower)
N_ant = 4
ant = flower$ant
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol = vol,
                  N_ant = N_ant,
                  ant = ant
)

write("// Stan model for simple total flowers regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
  }
}
model {
 // Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(mu);
}
",

"abort_ant_lm1.stan")
## Check that you wrote a file
stanc("abort_ant_lm1.stan")
## Save the file path
stanny <- "abort_ant_lm1.stan"
## Now we fit the model
fit_abort_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_abort_ant

posterior_abort_ant <- as.data.frame(fit_abort_ant)
print(colnames(posterior_abort_ant))

lm_alpha_other <- summary(abort_model_ant)$coeff[1]  # the intercept
lm_alpha_crem <- summary(abort_model_ant)$coeff[3]
lm_alpha_liom <- summary(abort_model_ant)$coeff[4]
lm_alpha_vacant <- summary(abort_model_ant)$coeff[5]
lm_beta_other <- summary(abort_model_ant)$coeff[2]  # the slope
lm_beta_crem <- summary(abort_model_ant)$coeff[6]
lm_beta_liom <- summary(abort_model_ant)$coeff[7]
lm_beta_vacant <- summary(abort_model_ant)$coeff[8]
